\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-01-17 papers/methods/paper.tex
Updating some des and lowz data
Adding population
Paper update: population caption

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelOl.pdf}
	\end{center}
	\caption{Posterior surfaces for 200 realisations of supernova data with the $\Lambda$CDM model. Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ and $\Omega_\Lambda$ due to the severe degeneracy between the parameters.}
	\label{fig:simple_ol}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelW.pdf}
	\end{center}
	\caption{Posterior surfaces for 200 realisations of supernova data with the Flat $w$CDM model. The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing incorrect statistics due to the highly non-Gaussian nature of the posterior surface.}
	\label{fig:simple_w}
\end{figure}

We generate simple toy data to validate the basic premise of the model. For both DES and LowZ data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$. Masses are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included. Absolute magnitudes are drawn from $\mathcal{N}(-19.3, 0.1)$, true stretch values are drawn from $\mathcal{N}(0, 1)$ and true colour values are drawn from $\mathcal{N}^{\rm Skew}(0, 0.1, 2)$.

Independent observational errors of $0.04$, $0.2$, $0.03$ on $m_B$, $x_1$ and $c$ (following the mean uncertainty for DES SNANA simulations) are added to create the observables. The results are then passed through the selection effects, where each supernova is only selected based on $P(S|m_B)$, using a skew normal function for the LowZ supernovae and error function for the DES-like supernovae. We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES spectroscopic analysis.

@@

\begin{table}
	\centering
	\caption{Maximum likelihood cosmological parameters determined from stacking the surfaces of 200 fits to independent realisations of toy supernova data. As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae. Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$. No significant bias is detected in either cosmological model.}
	\label{tab:simple_model}
	\begin{tabular}{ccc}
		\hline
		Model & $\Omega_m$ & $w$ \\ 
		\hline
		Flat $\Lambda$CDM & $0.301\pm 0.016$ & -- \\ 
		Flat $w$CDM + $\Omega_m$ prior & $\left( 300.5^{+10.5}_{-9.8} \right) \times 10^{-3}$ & $-0.998^{+0.043}_{-0.047}$ \\ 
		\hline
	\end{tabular}
\end{table}

@@

\begin{table}
	\centering
	\caption{Tested population distributions, where the {\sk} LowZ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
	\label{tab:dist}
	\begin{tabular}{l|llll}
		\hline
		Model & $\langle x_1 \rangle$ & $\sigma_{x_1}$ &  $\langle c \rangle$ & $\sigma_c$  \\
		\hline
		Gaussian LowZ &  $0.0$ & $1.0$ & $0.0$ & $0.1$ \\
		Gaussian DES  &  $0.0$ & $1.0$ & $0.0$ & $0.1$  \\
		{\sk} LowZ    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\
		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$ \\
		\hline
	\end{tabular}
\end{table}

@@

\begin{table}
	\centering
	\caption{Parameter summaries shown for 150 realisations of supernovae data, each comprising of 300 LowZ supernovae and 250 DES-like supernovae. Maximum likelihood statistics, with errors quoted to $1\sigma$ (68\% confidence) are reported.}
	\label{tab:bulk_summary}
	\begin{tabular}{ccccc}
		\hline
		Model & $\Omega_m$ & $\alpha$ & $\beta$  \\ 
		\hline
		C11 Gaussian & $0.316^{+0.062}_{-0.064}$ & $0.143^{+0.028}_{-0.027}$ & $2.97^{+0.20}_{-0.19}$  \\ 
		C11 {\sk} & $0.316^{+0.063}_{-0.062}$ & $0.155\pm 0.024$ & $2.77^{+0.26}_{-0.27}$  \\ 
		Combined Gaussian & $0.299^{+0.067}_{-0.060}$ & $0.144^{+0.030}_{-0.028}$ & $3.11^{+0.21}_{-0.22}$  \\ 
		Combined {\sk} & $0.297^{+0.075}_{-0.064}$ & $0.150^{+0.026}_{-0.025}$ & $3.00^{+0.29}_{-0.31}$  \\ 
		G10 Gaussian & $0.290^{+0.070}_{-0.060}$ & $0.145\pm 0.030$ & $3.27\pm 0.23$ \\ 
		G10 {\sk} & $0.290^{+0.079}_{-0.073}$ & $0.147^{+0.026}_{-0.028}$ & $3.26^{+0.34}_{-0.33}$ \\ 
		\hline
	\end{tabular}
\end{table}

>>>>

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{plot_pop_simple.pdf}
	\end{center}
	\caption{Population distributions shown in redshift and uncorrected absolute magnitude $m_B - \mu$ for 1000 supernovae in both high-redshift and low-redshift surveys. Selection effects are visible in both samples, where red supernovae are often cut as redshift increases.}
	\label{fig:simple_pop}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelOl.pdf}
	\end{center}
	\caption{Posterior surfaces for 200 realisations of supernova data with the $\Lambda$CDM model. Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ and $\Omega_\Lambda$ due to the severe degeneracy between the parameters.}
	\label{fig:simple_ol}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelW.pdf}
	\end{center}
	\caption{Posterior surfaces for 200 realisations of supernova data with the Flat $w$CDM model. The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing incorrect statistics due to the highly non-Gaussian nature of the posterior surface.}
	\label{fig:simple_w}
\end{figure}

We generate simple toy data to validate the basic premise of the model. For both high-redshift and low-redshift (LowZ) data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$. Masses are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included. Absolute magnitudes are drawn from $\mathcal{N}(-19.3, 0.1)$, true stretch values are drawn from $\mathcal{N}(0, 1)$ and true colour values are drawn from $\mathcal{N}^{\rm Skew}(0, 0.1, 2)$. Redshifts are drawn from a power law to increase the number of events as redshift increases.

Independent observational errors of $0.04$, $0.2$, $0.03$ on $m_B$, $x_1$ and $c$ (following the mean uncertainty for DES SNANA simulations) are added to create the observables. The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters. The results are then passed through the selection effects, where each supernova is only selected based on $P(S|m_B)$, using a skew normal function for the LowZ supernovae and error function for the DES-like supernovae. We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES spectroscopic analysis. Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, showing presence of strong selection effects in both toy surveys.

@@

\begin{table}
	\centering
	\caption{Maximum likelihood cosmological parameters determined from stacking the surfaces of 200 fits to independent realisations of toy supernova data. As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae. Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$. No significant bias is detected in either cosmological model.}
	\label{tab:simple_model}
	\begin{tabular}{l|cc}
		\hline
		Model & $\Omega_m$ & $w$ \\ 
		\hline
		Flat $\Lambda$CDM & $0.301\pm 0.016$ & -- \\ 
		Flat $w$CDM + $\Omega_m$ prior & $\left( 300.5^{+10.5}_{-9.8} \right) \times 10^{-3}$ & $-0.998^{+0.043}_{-0.047}$ \\ 
		\hline
	\end{tabular}
\end{table}

@@

\begin{table}
	\centering
	\caption{Tested population distributions, where the {\sk} LowZ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
	\label{tab:dist}
	\begin{tabular}{l|cccc}
		\hline
		Model & $\langle x_1 \rangle$ & $\sigma_{x_1}$ &  $\langle c \rangle$ & $\sigma_c$  \\
		\hline
		Gaussian LowZ &  $0.0$ & $1.0$ & $0.0$ & $0.1$ \\
		Gaussian DES  &  $0.0$ & $1.0$ & $0.0$ & $0.1$  \\
		{\sk} LowZ    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\
		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$ \\
		\hline
	\end{tabular}
\end{table}

@@

\begin{table}
	\centering
	\caption{Parameter summaries shown for 150 realisations of supernovae data, each comprising of 300 LowZ supernovae and 250 DES-like supernovae. Maximum likelihood statistics, with errors quoted to $1\sigma$ (68\% confidence) are reported.}
	\label{tab:bulk_summary}
	\begin{tabular}{lcccc}
		\hline
		Model & $\Omega_m$ & $\alpha$ & $\beta$  \\ 
		\hline
		C11 Gaussian & $0.316^{+0.062}_{-0.064}$ & $0.143^{+0.028}_{-0.027}$ & $2.97^{+0.20}_{-0.19}$  \\ 
		C11 {\sk} & $0.316^{+0.063}_{-0.062}$ & $0.155\pm 0.024$ & $2.77^{+0.26}_{-0.27}$  \\ 
		Combined Gaussian & $0.299^{+0.067}_{-0.060}$ & $0.144^{+0.030}_{-0.028}$ & $3.11^{+0.21}_{-0.22}$  \\ 
		Combined {\sk} & $0.297^{+0.075}_{-0.064}$ & $0.150^{+0.026}_{-0.025}$ & $3.00^{+0.29}_{-0.31}$  \\ 
		G10 Gaussian & $0.290^{+0.070}_{-0.060}$ & $0.145\pm 0.030$ & $3.27\pm 0.23$ \\ 
		G10 {\sk} & $0.290^{+0.079}_{-0.073}$ & $0.147^{+0.026}_{-0.028}$ & $3.26^{+0.34}_{-0.33}$ \\ 
		\hline
	\end{tabular}
\end{table}
